Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects

### 2015-03-10 inband-STN.tex
small fixes to algs
sync interface

\newcommand{\hide}[1]{}
\algblock[<block>]{<start>}{<end>}
\algblockdefx[Transaction]{startTransaction}{endTransaction} %
[0]{ \textbf{start transaction}} %
[1][res]{ $#1\gets$  \textbf{commit transaction}}

@@

\section{The Container Object}\label{sec:vision}

The container synchronization object introduced in this paper
is essentially a transactional memory abstraction which can be implemented
in standard Openflow switches.
Before presenting the technical details in the next section,
we give a brief overview.

The postulated container object can store an arbitrary number of \emph{items}.
As we will see, these items correspond to flow entries on the Openflow switch.
The same item may even exist multiple times in the container (i.e., the container
is a multi-set).
The container
provides two main functions:
\begin{enumerate}
\item \textbf{\emph{Conditional Multi-Add($S$):}} A set $S$ of items $s\in S$
is (atomically) added to the container, if and only if no element $s\in S$
already existed in the container. If such an element already existed,
the operation does not take effect (i.e., the transaction is aborted).

\item \textbf{\emph{Conditional Multi-Remove($S$):}} A set $S$ of items $s\in S$
is (atomically) removed from the container, if and only if each element $s\in S$
already existed in the container. If one or more elements did not exist,
the operation does not take effect (i.e., the transaction is aborted).
\end{enumerate}

The container object allows controllers to atomically and
concurrently read and modify state. This
allows us, for example, to solve consensus among any
number of controllers.
More generally, the container object
allows us to
achieve consistent network management.

\section{Openflow Implementation}\label{sec:container}

@@

\begin{algorithm}[t]
    \caption{Advanced Update Algorithm}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$, multi-set (with optional add validation) $Next$, set (with full validation) $Cur$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ unused \emph{id} (in both $Cur$ and $Next$)
 			\State $pid\gets$ current policy id (only id in $Cur$)
 			\State $Next$\emph{.add}($pid$)
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
 				\State $Next$\emph{.add}($my\_id$, True)
 				\State $Next$\emph{.delete}($my\_id$, True)
	 			\State DeleteGroup($pid$)
	 			\State AddGroup($my\_id$)
				\State $Next$\emph{.delete}($pid$)
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=\texttt{true}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\begin{algorithm}[t]
    \caption{\textbf{\emph{Multi-Add()}}}
    \label{alg:multi-add}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
%    \Ensure
    		\State $value \gets i$
    		\State $mask \gets all-1$
    		\State $cookie \gets i.caller$
    		\State $action \gets i.caller$
    		\If {$b\_unique$}
    			\State $flag \gets 0$
    		\Else
    		    \State $flag \gets CHECK\_OVERLAP$
    		\EndIf
    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Multi-Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $value \gets C.i$
    		\State $mask \gets all-1$
    		\State $cookie \gets C.i.caller$
    		\If {$b\_unique$}
    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
    		\Else
    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
    		\EndIf

@@

\begin{algorithm}[t]
    \caption{\textbf{\emph{Add()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require  id $i$.
%    \Ensure
    		\State $command \gets OFPGC\_ADD$
    		\State $type \gets OFPGT\_INDIRECT$
    		\State $group_id \gets i$
    		\State $buckets \gets [dummy\_backet]$
    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $command \gets OFPGC\_DELETE$
    		\State $group_id \gets i$
    		\State $cmd\gets GroupMod(command, group\_id) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

>>>>

\newcommand{\hide}[1]{}
\newcommand{\concat}[0]{\frown}
\newcommand{\append}[0]{\frown}

\algblock[<block>]{<start>}{<end>}
\algblockdefx[Transaction]{startTransaction}{endTransaction} %
[0]{ \textbf{start transaction}} %
[1][res]{ $#1\gets$  \textbf{commit transaction}}

@@

%\section{The Container Object}\label{sec:vision}
\section{Transactional synchronization}\label{sec:trans}

In this section, we sketch the synchronization abstraction we build on
top of standard OpenFlow functionality. At a high level, we export a
transactional interface which allows a controller to perform multiple
operations on the configuration of an OpenFlow switch as an atomic
\emph{transaction}, inspired by the concept of transactional
memory~\cite{stm-st95,tm-book}. A transaction may \emph{commit}, in
which case it appears as executed sequentially, or \emph{abort}, in
which none of its operations takes effect.

\subsection{Interface}\label{sec:t-if}
We assume that transactions may comprise operations on the flow table
of an SDN switch, such as reading flow tables or adding new rules. 
 In addition, we also assume transactional access to  \emph{multi-set}
abstractions that store \emph{items} and export \emph{add} and \emph{delete} operations with
regular semantics. The multi-set abstractions are used for resolving
conflicts between concurrent updates of the switch configuration and are themselves 
implemented as flow-table components.

A multi-set object also exports operations \emph{ms-add} and \emph{ms-delete} with the following
semantics:

%The container synchronization object introduced in this paper
%is essentially a transactional memory abstraction which can be implemented
%in standard Openflow switches.
%Before presenting the technical details in the next section,%
%we give a brief overview.

%The postulated container object can store an arbitrary number of \emph{items}.
%As we will see, these items correspond to flow entries on the Openflow switch.
%The same item may even exist multiple times in the container (i.e., the container
%is a multi-set).
%The container
%provides two main functions:
\begin{itemize}
\item \textbf{$\textit{cond-add}(x)$:} if $x$ is not in the set, adds 
  $x$ to the set; otherwise, \emph{fails} and, respectively,
  aborts the transaction in which the operation is invoked.

\item \textbf{$\textit{cond-delete}(x)$:} if $x$ is in the set, deletes an
$x$ from the set; otherwise, \emph{fails} and
  aborts the transaction.
\end{itemize}

Formally, we guarantee that in every execution, all committed
transactions constitute a \emph{legal} sequential
order~\cite{Pap79-serial}. On the progress side, we ensure that a
transaction $T$ is only allowed to abort either because another 
transaction updated a flow table entry read by $T$  
or if one of $T$'s multi-set operations fails.

Note that
%, assuming such a transactional memory,  
a $\emph{contains}(x)$ operation can be implemented as a transaction
composed of $\textit{cond-add}(x)$ followed by $\textit{cond-delete}(x)$.  
A \emph{set} abstraction is a special instance of a multi-set that is
only modified with \textit{cond-add} and \textit{cond-delete}
operations (this way the set only contains at most one copy of a given
item).

\subsection{Transactions in Openflow}\label{sec:t-impl}

@@

\begin{algorithm}[t]
    \caption{Advanced Update Algorithm}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$, multi-set (with optional add validation) $Next$, set (with full validation) $Cur$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ unused \emph{id} (in both $Cur$ and $Next$)
 			\State $pid\gets$ current policy id (only id in $Cur$)
 			\State $Next$\emph{.add}($pid$)
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
 				\State $Next$\emph{.add}($my\_id$, True)
 				\State $Next$\emph{.delete}($my\_id$, True)
	 			\State CUR\emph{.delete}($pid$)
	 			\State CUR\emph{.add}($my\_id$)
				\State $Next$\emph{.delete}($pid$)
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=\texttt{true}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

\begin{algorithm}[t]
    \caption{\textbf{\emph{MultiSet-Add()}}}
    \label{alg:multi-add}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
%    \Ensure
    		\State $value \gets i\concat i$
    		\State $mask \gets caller\concat all-1$
    		\State $cookie \gets i\concat caller$
    		\State $action \gets i\concat caller$
    		\If {$b\_unique$}
    			\State $flag \gets 0$
    		\Else
    		    \State $flag \gets CHECK\_OVERLAP$
    		\EndIf
    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{MultiSet-Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $value \gets C.i$
    		\State $mask \gets all-1$
    		\State $cookie \gets C.i.caller$
    		\If {$b\_unique$}
    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
    		\Else
    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
    		\EndIf

@@

\begin{algorithm}[t]
    \caption{\textbf{\emph{Set-Add()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require  id $i$.
%    \Ensure
    		\State $command \gets OFPGC\_ADD$
    		\State $type \gets OFPGT\_INDIRECT$
    		\State $group_id \gets i$
    		\State $buckets \gets [dummy\_backet]$
    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Set-Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $command \gets OFPGC\_DELETE$
    		\State $group_id \gets i$
    		\State $cmd\gets GroupMod(command, group\_id) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}
